Microclimatic spatial variation buffers local warming in alpine vegetation

### 2022-01-20 doc/manuscript.Rmd
new nmds
species chorology

@@

The study was conducted in the calcareous massif of Picos de Europa National Park, in northern Spain (Figure 1A). The study area is a biodiversity hotspot for cold-adapted plants in the Iberian Peninsula and a biogeographical hub for Alpine and Mediterranean lineages in Western Europe (XXX). The main calcareous massif occupies c. 50 km2 and support high diversity of ecosystems, with alpine vegetation mostly occurring between 1900 and 2400 m a.s.l., with a local species pool of XX species (Jiménez-Alfaro et al. XXX). In 2008, we established a long-term ecological research program for monitoring soil climate and vegetation change. We selected four study sites along a north-south gradient, reflecting the main macroclimatic variation from oceanic to Mediterranean influence. In 2016, we additionally installed a Water potential logger (XXX) to characterize water stress in the four sites and to .... The sites were selected to... general design and description of sites... (Borja)

> Empiezo a pensar que sería mejor quitar esto, por mucho que me fastidie, y guardarnos el tema del WP para otro paper junto con lo de Omaña. Eso simplificaría el gráfico (de otro modo, tendría más lógica ponerlo con la Figura 1, porque en realidad no hacemos nada con esto)

@@

In the temporal survey (2 plots per site, n =16), we recorded XX species in 2008 and XX in 2018. (species loss, gains)... Changes in species frequency were/were not correlated with XX and XX.... (Table 1) (Eduardo). In the spatial survey (20 plots per site, n = 80; including four plots placed in rocks with no plants) we recorded XX species. Changes in species composition were associated to GDD (NMDS axis 1) to represent a main gradient from the warmest to the coldest sites (Figure 4), and to FDD (NMDS2) in a gradient of snow versus frost winters. In the compositional space, the sites were poorly differentiated across the four sites (PERMANOVA, adonis) mainly by the effect of one site (Los Boches) with most plots places in the cold and frost margins. Etc.... (Borja).

>>>>

```{r}
library(vegan)
load("../results/numerical/environmental-fit.R")
load("../results/numerical/bioclim-correlations.R")
```

@@

The study was conducted in the calcareous massif of Picos de Europa National Park, in northern Spain (Figure 1A). The study area is a biodiversity hotspot for cold-adapted plants in the Iberian Peninsula and a biogeographical hub for Alpine and Mediterranean lineages in Western Europe (XXX). The main calcareous massif occupies c. 50 km^2^ and support high diversity of ecosystems, with alpine vegetation mostly occurring between 1900 and 2400 m a.s.l., with a local species pool of XX species (Jiménez-Alfaro et al. XXX). In 2008, we established a long-term ecological research program for monitoring soil climate and vegetation change. We selected four study sites along a north-south gradient, reflecting the main macroclimatic variation from oceanic to Mediterranean influence. The sites were selected to... general design and description of sites... (Borja)

@@

In the temporal survey (2 visits x 2 plots per site, n = 16) we recorded 42 species in 2009 and 45 in 2019. Of the species recorded in 2009, we did not find again in 2019 the following three: *Festuca burnatii*, *Galium pyrenaicum* and *Viola riviniana*. Conversely, in 2019 we recorded six species that we had not seen in 2009: *Arenaria purpurascens*, *Gentiana verna*, *Potentilla crantzii*, *Sedum brevifolium* and *Seseli montanum*. Changes in species frequency were/were not correlated with XX and XX.... (Table 1) (Eduardo). In the spatial survey (20 plots per site, n = 80; including one plot placed in rocks with no vascular plants) we recorded 81 species, 37 of which were absent from the temporal survey plots. Across the whole study system, the five most frequent species were *Thymus praecox* subsp. *ligusticus* (83 occurences), *Anthyllis vulneraria* (73), *Koeleria vallesiana* (59), *Minuartia verna* (55) and *Helianthemum canum* (52).

Changes in species composition (**Fig. 4**) were associated to GDD (NMDS axis 1) to represent a main gradient from the warmest to the coldest microsites, and to FDD (NMDS2) in a gradient of snow versus frost winters. In the compositional space, the sites were poorly differentiated across the four sites (PERMANOVA, adonis) mainly by the effect of one site (Los Boches) with most plots places in the cold and frost margins. Etc.... (Borja).

Results of the environmental fit:

```{r}
print(ef1)
```

Correlations between bioclimatic variables:

```{r}
print(biocor)
```

Let us keep GDD, bio7 and FDD for plotting NMDS (low autocorrelation, high r2 in environmental fit)
